\textsc{HubbleCLIP}: Associating Astronomical Observations and Natural Language with Multi-Modal Models

### 2024-01-02 paper/main.tex
Add some comments

\newcommand{\hubble}{\emph{Hubble}\xspace}

@@

@@

@@

We curate a dataset of images of \hubble observations and corresponding text descriptions. We rely on summarized versions of proposal abstracts from the Proposal Abstracts Catalog\footnote{\url{https://archive.stsci.edu/hst/proposal_abstracts.html}} -- a catalog of all accepted \hubble proposals -- as captions for the observations. The HST has been operational for 33 years, having been launched on April 24, 1990, and we use proposals and observations up to Cycle 31.

@@

@@

@@

\begin{landscape}
      \begin{table}[h!]
          \centering
          \begin{tabular}{m{1.8cm} m{8cm} m{5cm} m{6cm}}
              \toprule
              \bfseries Prop. ID & \centering\arraybackslash \bfseries Proposal abstract & \multicolumn{2}{c}{\bfseries LLM-extracted summary} \tabularnewline
              \cmidrule(r){3-4}
              & & \centering\arraybackslash \bfseries Objects and phenomena & \centering\arraybackslash \bfseries Science use cases \tabularnewline
              \midrule
              \input{\thedatafolder/id1_0.txt} & {\scriptsize \input{\thedatafolder/abs1_0.txt}} & {\scriptsize \input{\thedatafolder/obj1_0.txt}} & {\scriptsize \input{\thedatafolder/sci1_0.txt}} \tabularnewline
              \midrule
              \input{\thedatafolder/id1_1.txt} & {\scriptsize \input{\thedatafolder/abs1_1.txt}} & {\scriptsize \input{\thedatafolder/obj1_1.txt}} & {\scriptsize \input{\thedatafolder/sci1_1.txt}} \tabularnewline
              \midrule
              \input{\thedatafolder/id1_2.txt} & {\scriptsize \input{\thedatafolder/abs1_2.txt}} & {\scriptsize \input{\thedatafolder/obj1_2.txt}} & {\scriptsize \input{\thedatafolder/sci1_2.txt}} \tabularnewline
              \midrule
              \input{\thedatafolder/id1_3.txt} & {\scriptsize \input{\thedatafolder/abs1_3.txt}} & {\scriptsize \input{\thedatafolder/obj1_3.txt}} & {\scriptsize \input{\thedatafolder/sci1_3.txt}} \tabularnewline
              \bottomrule
          \end{tabular}
          \caption{Examples of the beginning of raw proposal abstracts (second column) and LLM (\textsc{Mixtral-8x7B})-extracted summaries (rightmost two columns), separately extracting objects and phenomena as well as potential downstream science use cases. The LLM-extracted summaries are used for associating text with observations.}
          \label{tab:datasetsumm}
      \end{table}
  \end{landscape}

@@

CLIP \citep[Contrastive Language-Image Pretraining;][]{radford2021learning} is a multi-modal model pre-trained on a large corpus of image-text pairs via weak supervision using a contrastive loss. Given a minibatch $\mathcal{B}$ of $|\mathcal{B}|$ image-text pairs $\{(I_i, T_i)\}$, the goal is to align the learned representations of corresponding (positive) pairs $(I_i, T_i)$ while repelling the representations of unaligned (negative) pairs $(I_i, T_{j\neq i})$. Image and text encoders $f(\cdot)$ and $g(\cdot)$ are used to map images and text to a common embedding space, and the standard softmax-based InfoNCE~\citep{oord2018representation} contrastive loss function is given by
\begin{equation}
\mathcal L(\mathcal B) = -\frac{1}{2|\mathcal{B}|} \sum_{i=1}^{|\mathcal{B}|}\left(\overbrace{\log \frac{e^{t {x}_i \cdot {y}_i}}{\sum_{j=1}^{|\mathcal{B}|} e^{t {x}_i \cdot {y}_j}}}^{\text {image } \rightarrow \text { text softmax }}+\overbrace{\log \frac{e^{t {x}_i \cdot {y}_i}}{\sum_{j=1}^{|\mathcal{B}|} e^{t {x}_j \cdot {y}_i}}}^{\text {text } \rightarrow \text { image softmax }}\right)
\end{equation}
where ${x}_i=\frac{f\left(I_i\right)}{\left\|f\left(I_i\right)\right\|_2}$ and ${y}_i=\frac{g\left(T_i\right)}{\left\|g\left(T_i\right)\right\|_2}$ are the normalized representations of the $i$-th image and text, respectively, and $t$ is a (typically learnable) temperature hyperparameter.

@@

The base CLIP model is fine-tuned using the dataset described in Sec.~\ref{sec:dataset}, using either the LLM-summarized data or the raw proposal abstracts. When using raw proposal abstracts, random chunks of the text delimited by periods are selected on the fly to fit within the maximum length of the text encoder. Images are downsampled to the native resolution of the image encoder and randomly rotated over training. Given the relatively modest size of the fine-tuning set, a batch size $|\mathcal B| = 32$ is used throughout. We note that positive and negative image-text association is noisy and imperfect, since multiple images can be associated with the same abstract.

@@

@@

>>>>

\newcommand{\package}[1]{\textsl{#1}\xspace}
\newcommand{\eg}{{e.\,g.}\xspace}
\newcommand{\ie}{{i.\,e.}\xspace}
\newcommand{\SM}[1]{\textcolor{blue}{[SM: #1]}}
\newcommand{\hubble}{\emph{Hubble}\xspace}

\def\preprintno{XXXX} % Insert correct preprint number

@@

% Define a new fancy page style
\fancypagestyle{firstpage}{
    \rhead{MIT-CTP/\preprintno}
    % Define other header and footer elements if necessary
}

@@

\thispagestyle{firstpage}

@@

We curate a dataset of images of \hubble observations and corresponding text descriptions. We rely on summarized versions of proposal abstracts from the Proposal Abstracts Catalog\footnote{\url{https://archive.stsci.edu/hst/proposal_abstracts.html}} -- a catalog of all accepted \hubble proposals -- to derive captions for the observations. The HST has been operational for 33 years, having been launched on April 24, 1990. We use available proposals and observations up to Cycle 30, which commenced data-taking in 2022.

@@

% Mention that the signal is noisy!

@@

% Note that the signal is very imperfect!

@@

\begin{landscape}
      \begin{table}[h!]
          \centering
          \begin{tabular}{m{1.8cm} m{8cm} m{5cm} m{6cm}}
              \toprule
              \bfseries Prop. ID & \centering\arraybackslash \bfseries Proposal abstract & \multicolumn{2}{c}{\bfseries LLM-extracted summary} \tabularnewline
              \cmidrule(r){3-4}
              & & \centering\arraybackslash \bfseries Objects and phenomena & \centering\arraybackslash \bfseries Science use cases \tabularnewline
              \midrule
              \input{\thedatafolder/id1_0.txt} & {\scriptsize \input{\thedatafolder/abs1_0.txt}} & {\scriptsize \input{\thedatafolder/obj1_0.txt}} & {\scriptsize \input{\thedatafolder/sci1_0.txt}} \tabularnewline
              \midrule
              \input{\thedatafolder/id1_1.txt} & {\scriptsize \input{\thedatafolder/abs1_1.txt}} & {\scriptsize \input{\thedatafolder/obj1_1.txt}} & {\scriptsize \input{\thedatafolder/sci1_1.txt}} \tabularnewline
              \midrule
              \input{\thedatafolder/id1_2.txt} & {\scriptsize \input{\thedatafolder/abs1_2.txt}} & {\scriptsize \input{\thedatafolder/obj1_2.txt}} & {\scriptsize \input{\thedatafolder/sci1_2.txt}} \tabularnewline
              \midrule
              \input{\thedatafolder/id1_3.txt} & {\scriptsize \input{\thedatafolder/abs1_3.txt}} & {\scriptsize \input{\thedatafolder/obj1_3.txt}} & {\scriptsize \input{\thedatafolder/sci1_3.txt}} \tabularnewline
              \bottomrule
          \end{tabular}
          \caption{Examples of the initial parts of raw proposal abstracts (second column) and LLM (\textsc{Mixtral-8x7B})-extracted summaries (rightmost two columns), separately extracting objects and phenomena as well as potential downstream science use cases. The LLM-extracted summaries are used for associating text with observations.}
          \label{tab:datasetsumm}
      \end{table}
  \end{landscape}

@@

CLIP \citep[Contrastive Language-Image Pretraining;][]{radford2021learning} is a multi-modal model pre-trained on a large corpus of image-text pairs via weak supervision using a contrastive loss. Given a minibatch $\mathcal{B}$ of $|\mathcal{B}|$ image-text pairs $\{(I_i, T_i)\}$, the goal is to align the learned representations of corresponding (positive) pairs $(I_i, T_i)$ while repelling the representations of unaligned (negative) pairs $(I_i, T_{j\neq i})$. Image and text encoders $f(\cdot)$ and $g(\cdot)$ are used to map images and text to a common embedding space. The standard softmax-based bidirectional variation of the InfoNCE~\citep{oord2018representation} contrastive loss function, as used by CLIP, is particularly effective for multimodal learning. This bidirectionality is crucial as it ensures the model learns to map both images to text and text to images with equal importance. This symmetry in learning is essential for tasks that require a mutual understanding and interchangeability between visual and textual representations, such as image captioning, text-to-image synthesis, and cross-modal retrieval. The bidirectional loss is given by \citep{radford2021learning}
\begin{equation}
\mathcal L(\mathcal B) = -\frac{1}{2|\mathcal{B}|} \sum_{i=1}^{|\mathcal{B}|}\left(\overbrace{\log \frac{e^{{x}_i \cdot {y}_i / \tau}}{\sum_{j=1}^{|\mathcal{B}|} e^{{x}_i \cdot {y}_j / \tau}}}^{\text {image } \rightarrow \text { text softmax }}+\overbrace{\log \frac{e^{{x}_i \cdot {y}_i / \tau}}{\sum_{j=1}^{|\mathcal{B}|} e^{{x}_j \cdot {y}_i / \tau}}}^{\text {text } \rightarrow \text { image softmax }}\right)
\end{equation}
where ${x}_i=\frac{f\left(I_i\right)}{\left\|f\left(I_i\right)\right\|_2}$ and ${y}_i=\frac{g\left(T_i\right)}{\left\|g\left(T_i\right)\right\|_2}$ are the normalized representations of the $i$-th image and text, respectively, and $\tau$ is a (typically learnable) temperature hyperparameter. This bidirectional approach not only enhances the model's ability to create robust and versatile embeddings but also aligns with the intuitive understanding of how humans interact with and interpret multimodal information, further bridging the gap between artificial and natural intelligence in multimodal contexts.

@@

The base CLIP model is fine-tuned using the dataset described in Sec.~\ref{sec:dataset}, using either the LLM-summarized data or the raw proposal abstracts. When using raw proposal abstracts, random chunks of the text delimited by periods are selected on the fly to fit within the maximum token length of the text encoder. Images are randomly cropped to the native resolution of the image encoder and randomly rotated at each training step. Given the relatively modest size of the fine-tuning set, a batch size $|\mathcal B| = 32$ is used throughout. We note that the positive and negative image-text association is noisy and imperfect, since multiple images can be associated with the same abstract.

@@

% Should write BIS!

@@

% Code repo and package cites.
% Dataset availability
